### 2022-05-31 main/chapter_2ljets.tex
expan ewk with problems

\subsection{Electroweak}
\label{sec:2ljets_origins_electroweak}
% susy motivation

As motivated in the previous sections, I have personal tastes for how searches
should be constructed.
One taste is for simulation-driven background estimates, since they are
interpretable as representations of the Standard Model, and they can exceed the
precision available in data alone.
Another is for orthogonal regions, such that their likelihoods are independent
and the information in their data can be cleanly combined.

The core idea behind the design of the $\twoljets$-electroweak analysis is to
update the regions of the partial Run~2 search in its $\twoljets$
channel~\cite{atlas_23l_SUSY_2016_24}, which also searched for signal models
producing chargino-neutralino pairs which decay via $W$ and $Z$ bosons.
A particular goal was to improve sensitivity by replacing their overlapping
`SR2-int' and `SR2-high' regions with orthogonal, combined selections.
\emph{I am grateful to Sarah Williams for suggesting this approach.}

As in the $\twoljets$-strong search, the partial Run~2 search chose one
$\twoljets$ signal region from an expected sensitivity for each signal point.
Two regions and the exclusion contours from that search are displayed in
Figure~\ref{fig:2ljets_23l_stuff}.

@@

Additionally, $\twoljets$ widens its scope to target a simplified GMSB model,
which has been illustrated in Figure~\ref{fig:2ljets_signal_diagrams}.
These two models behave similarly since both feature decays via massive bosons
($WZ$, $ZZ$ or $Zh$) to leptons and jets, in similarly structures decay trees
to pairs of invisible supersymmetric particles.
Although widening sensitivity to both models does sacrifice sensitivity to
either one, the increased degree of model independence is itself of value for
future reinterpretations of the data.

@@

\subsection{\texorpdfstring{$\met$ significance}{Etmiss significance}}
\label{sec:2ljets_metsig}
Most common Standard Model processes at the LHC do not produce hard,
invisible particles.
Our supersymmetric signals, however, do.
This makes the missing transverse momentum $\ptmiss$ its magnitude $\met$
important features for separating signal and background events in all parts
of the $\twoljets$ search.

>>>>

\subsection{Electroweak}
\label{sec:2ljets_origins_electroweak}
As motivated in the previous sections, I have personal preferences for
simulation-driven backgrounds and orthogonal multi-bin regions.
This is because simulation-driven estimates are interpretable as
Standard Model predictions,
and orthogonal regions have independent likelihoods which can be cleanly
combined to leverage the information in their data.

The main idea behind the design of the $\twoljets$-electroweak analysis is to
update the regions of the partial Run~2 search in its $\twoljets$
channel~\cite{atlas_23l_SUSY_2016_24} which also searched for signal models
producing chargino-neutralino pairs which decay via $W$ and $Z$ bosons.
A particular goal was to improve sensitivity by replacing their overlapping
`SR2-int' and `SR2-high' regions with orthogonal selections.
\emph{I am grateful to Sarah Williams for suggesting this approach.}

As in the $\twoljets$-strong search, the partial Run~2 search chose one
signal region from an expected sensitivity for each signal point,
as displayed in
Figure~\ref{fig:2ljets_23l_stuff}
with the two regions and exclusion contours.

@@

Additionally, $\twoljets$ widens its scope to target a GMSB simplified model
illustrated in Figure~\ref{fig:2ljets_signal_diagrams}.
The C1N2 and GMSB models behave similarly since both feature decays via massive
bosons ($WZ$, $ZZ$ or $Zh$) to leptons and jets, in similarly structures decay
trees to pairs of invisible supersymmetric particles.
Although widening sensitivity to both models does sacrifice sensitivity to
either one, the increased degree of model independence is itself of value for
future reinterpretations of the data.
We are happy to sacrifice some precision for generality.

Multi-bin fits are not special.
They are standard \atlas\ practice, and there could have been no complication
in replacing the overlapping SR2-int and SR2-high with orthogonally binned
regions.
However, my design decisions for $\twoljets$-electroweak
caused problems for the clarity of communication and accuracy of background
modelling.
These issues are developed further through the remainder of this chapter.

A multi-binned histogram is usually and clearly illustrated as slicing up one
variable.
My design of $\twoljets$-electroweak violates this.
Instead, it is designed as a tree whose branches recursively split various
variables.
This leads to specifications which are challenging to present and understand,
and which require region names with numerous suffixes to specify where they
live in that tree.

Worse, I disregarded the idea that every region should have a large
background expectation.
Normally, this is justified by the limitations of approximate statistical
methods, which work
``from as few as $\mathcal{O}(10)$ data events''~\cite{histfitter2014}.
But the Poisson likelihood works for infinitesimally small
expectations~\cite{skilling2009cosmology}, and in small-bin limit one has
continuous fits (sometimes called ``extended''~\cite{barlow1990extended}),
which are standard and easily analysed.

The real reason to avoid small backgrounds is, in this context, to allow honest
assignments of background uncertainties.
Although main simulated samples may have good precision,
many `systematic' variations and `data-driven' estimates are also required.
Many of these alternative estimates have far worse precision, and can fail
totally as the number of simulated events accepted plunges towards zero;
they can require $\approx10$ data events to work properly.

Unfortunately, $\twoljets$-electroweak selections were finalized before I
appreciated this problem that background variations were dubious.
Various workaround are invoked to assign plausible variations despite small
yields, which we shall in the remainder of this chapter.

As seen in the splash summary of Section~\ref{sec:2ljets_splash}, this search
finds agreement with backgrounds except for some local deficits of data below
backgrounds.
Exclusion contours exceed their expected range because of these deficits,
but they can only be trusted if the background model is trusted.

@@

\subsection{\texorpdfstring{$\met$}{ETmiss} significance}
\label{sec:2ljets_metsig}
Most common Standard Model processes at the LHC do not produce hard,
invisible particles.
Our supersymmetric signals, however, do.
This makes the missing transverse momentum $\ptmiss$ its magnitude $\met$
important features for separating signal and background events in all parts
of the $\twoljets$ search.
